\documentclass[draft]{llncs}
\usepackage[show]{ed}
\usepackage{calbf}
\usepackage{amstext,amsmath,amssymb}
\usepackage{xspace}

\usepackage{mdframed}
\newenvironment{boxedquote}{\begin{mdframed}[leftmargin=1cm,rightmargin=1cm]}{\end{mdframed}}

\usepackage{wrapfig,paralist}
\usepackage[hyperref,style=alphabetic]{biblatex}
\addbibresource{kbibs/kwarcpubs.bib}
\addbibresource{kbibs/extpubs.bib}
\addbibresource{kbibs/kwarccrossrefs.bib}
\addbibresource{kbibs/extcrossrefs.bib}
\addbibresource{rest.bib}

\pagestyle{plain}
\usepackage{tikz}\usetikzlibrary{mmt,fit}

\usepackage{hyperref}
\title{System Description: A Semantics-Aware {\LaTeX}-to-WORD/ODF Converter}
\author{Lukas Kohlhase and Michael Kohlhase}
\institute{
  Math/CS, Jacobs University Bremen}
\begin{document}
\maketitle
\begin{abstract}
  \ednote{write something}
\end{abstract}

\section{Problem \& State of the Art}\label{sec:intro}

Many researchers in STEM fields only use {\LaTeX} to typeset their documents. However many
people still use Microsoft Word/Open office exclusively for their typesetting. When these
two groups of people intersect, it can lead to friction, as transforming text to {\LaTeX}
is quite trivial but not the opposite. For example if a conference requested documents in
Word format, the only recouse is often to just write the document in Word, which is a
pain, especially if any Mathematics is to be included.

\ednote{Here we state the Problem, some conferences and admin want papers in word format, however LaTeX is superior for various reasons. Hence converter is needed. Two step process, wastes some time.}

There are several methods to transform papers from {\LaTeX} to Word. The first method is
to just generate a PDF file and then open this file in Word/Open Office. However this
approach has several problems. First of all it does not convert semantically at all. A
headline is not treated as a headline, but as a piece of centered text with larger
font. Additionally math is treated as either not distinct from text in any meaningful way,
or as an image, making later editing impractical if not impossible. The same applies to
references. Finally it is a two step process that would save some time if creating the pdf
could be skipped.  \ednote{make PDF then open it with Word. Treats Math as either
  formatted Text or as pictures, unsuited for searching or for editing the math
  afterwards. References are also just pretty printed Text} \ednote{Various other ways
  that start at the .tex file, all either are buggy/not updated anymore or can't deal with
  math at all.}

\section{Implementation}\label{sec:impl}
\ednote{First convert to Ltxml using LaTeXML}
\ednote{Then convert the resulting ltxml using xslt stylesheets to a word apppropriate format. }
\ednote{Finally use postprocessing to zip it all up}
\ednote{Don't know where to put an explanation of Word format in this place}

\section{Conclusion}\label{sec:concl}
\ednote{MK@MK: say something}
\ednote{In Conclusion, easy to use}
\printbibliography
\end{document}


