\parindent=2true cm
\def\Lit#1{\medskip\item{[#1]} }
\Lit{A} T. Aubin, 
           {\it Nonlinear Analysis on Monifolds. Monge-Ampere Equations},
           Springer Verlag, 1982.
\Lit{BC} C. BÑr,
           {\it Private Korrespondez}
\Lit{BP} P. H. B\'erard,
           {\it Spectral Geometry: Direct and Inverse Problems},
           Lecture Notes in Mathematics {\bf 1207}, Springer-Verlag, 1986.
\Lit{CC} C. Caratheodory
            {\it Variationsrechnung und Partielle Differentialgleichungen
            2. Ordnung}
            Teubner, 1935.
\Lit{CH} C. Chevalley
            {\it Theory of Lie Groups}
            Princeton University Press, 1946.
\Lit{GH} M. Giaguinta und S. Hildebrandt,
           {\it Calculus of Variations, Vol 1},
            Preprint, Bonn, 1988.
\Lit{GT} D. Gilbarg und N. Trudinger,
           {\it Elliptic Partial Differential Equations of Second Order},
           Springer Verlag, New York, 1977.
\Lit{HL} R. Harvey und H.B. Lawson,
           {\it Calibrated Geometries},
           Acta Math. {\bf 184} (1982), 47-157.
\Lit{HS} S. Hildebrandt,
            {\it Vorlesungsskriptum Variationsrechnung}
            Bonn, 1975.
\Lit{LB} H.B. Lawson,
           {\it Lectures on Minimal Submanifolds Vol.1},
           Publish or Perish INC. 1980.  
\Lit{LL} L. Lichtenstein,
           {\it Untersuchungen Åber zweidimsionale regulÑre 
                Variationsprobleme I},
           Monatshefte fÅr Mathematik und Physik {\bf 28} (1917), 3-51.         
\Lit{LX} X. Li,       
           {\it Eine hinreichende Bedingung fÅr Minima und 
                Verzweigungsprobleme fÅr mehrfache Variationsintegrale},
           Diplomarbeit (1985), Bonn.
\Lit{M}  C. B. Morrey,
           {\it Multiple Integrals in the Calculus of Variations},
           Springer Verlag, New York, 1966.
\Lit{P}  R. S. Palais,
           {\it Foundations of global non-linear analysis},
           W. A. Benjamin, INC., New York Amsterdam, 1968.
\Lit{SB} B. Solomon,
           {\it On Foliations of $\RR^{n+1}$ by minimal hypersurfaces}
           Comment. Math. Helvet. {\bf 61} (1986), 67-83.
\Lit{SL} L. Simon,
           {\it Lectures on Geometric Measure Theory}
           Proceedings of the Centre for Mathematical Analysis. 
           Australian National University. Volume {\bf 3}, 1983.
\Lit{SN} N. Smale,
           {\it A Bridge Principle for Minimal and Constant Mean Curvature
                Submanifolds of $\RR^N$},
           Inventiones Math. {\bf 90} (1987) 505-549.
\Lit{Y}  K. Yosida,
           {\it Functional Analysis},
           Grundlehren der mathematischen Wissenschaften in 
           Einzelderstellungen Band {\bf 123}, Springer Verlag, 1968.
\parindent=0pt
